An Algorithm For Spatial Mapping Using a Hexagon Tilegram, With Application to Australian Maps

### 2019-07-30 algorithmRjournal/algorithmRjournal.tex
Cleaned file

There are four tasks to create a hexagon tilegram. These steps can be
executed by the main function, \texttt{create\_hexmap}, or can be
applied separately for more flexibility.

@@

\begin{Schunk}
\begin{Sinput}
# To install sugarbag:
# devtools::install_github("srkobakian/sugarbag")
library(sugarbag)

library(tidyverse)
\end{Sinput}
\begin{Soutput}
#> Registered S3 methods overwritten by 'ggplot2':
#>   method         from 
#>   [.quosures     rlang
#>   c.quosures     rlang
#>   print.quosures rlang
\end{Soutput}
\begin{Soutput}
#> Registered S3 method overwritten by 'rvest':
#>   method            from
#>   read_xml.response xml2
\end{Soutput}
\begin{Soutput}
#> -- Attaching packages --------------------------------------------------- tidyverse 1.2.1 --
\end{Soutput}
\begin{Soutput}
#> v ggplot2 3.1.1       v purrr   0.3.2  
#> v tibble  2.1.1       v dplyr   0.8.0.1
#> v tidyr   0.8.3       v stringr 1.4.0  
#> v readr   1.3.1       v forcats 0.4.0
\end{Soutput}
\begin{Soutput}
#> -- Conflicts ------------------------------------------------------ tidyverse_conflicts() --
#> x dplyr::filter() masks stats::filter()
#> x dplyr::lag()    masks stats::lag()
\end{Soutput}
\end{Schunk}

@@

The \texttt{create\_hexmap} function requires several parameters, if
they are not provided, the information will be derived from the polygon
set used. If users already have a grid and set of centroids they may
choose to only use the \texttt{allocate} function.

\textbf{The following must be provided to \texttt{create\_hexmap}:}

@@

\begin{Schunk}
\begin{Sinput}
## Use the create_centroids function to find polygon centroids:
centroids <- create_centroids(shp_sf = tas_lga, sf_id = "LGA_CODE16")
\end{Sinput}
\begin{Soutput}
#> Warning in st_centroid.sf(.): st_centroid assumes attributes are constant
#> over geometries of x
\end{Soutput}
\begin{Soutput}
#> Warning in st_centroid.sfc(st_geometry(x), of_largest_polygon =
#> of_largest_polygon): st_centroid does not give correct centroids for
#> longitude/latitude data
\end{Soutput}
\begin{Sinput}
## Join to data for each area
centroids <- left_join(centroids,
  sf::st_drop_geometry(tas_lga), by = "LGA_CODE16")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
fort_lga <- fortify_sfc(tas_lga)
polys <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") +theme_void() + coord_equal()

cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroids) + theme_void() + coord_equal()

gridExtra::grid.arrange(polys, cents, nrow = 1)
\end{Sinput}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/end_cents-1} \caption[Local Government Areas and derived centroids]{Local Government Areas and derived centroids.}\label{fig:end_cents}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
## create the bounding box from the set of polygon centroids:
bbox <- tibble::tibble(min = c(min(centroids$longitude, na.rm = TRUE), min(centroids$latitude, na.rm = TRUE)),
        max = c(max(centroids$longitude, na.rm = TRUE), max(centroids$latitude, na.rm = TRUE)))

hex_size = 0.3

# default size calculation:
#(bbox$max[1] - bbox$min[1])/(bbox$max[2] - bbox$min[2]) / 5

buffer_dist <- max((bbox$max[1] - bbox$min[1]), (bbox$max[2] - bbox$min[2]))*0.3
\end{Sinput}
\end{Schunk}

@@

A sequence is created for longitude columns, and another for latitude
rows.

The sequences begin at the minimum longitude (latitude), minus the
buffer distance. Equally spaced intervals the size of the hexagons are
created up to the maximum longitude (latitude), plus the buffer
distance.

An individual point is created from all intersections of the longitude
columns and latitude row sequences.

\begin{Schunk}
\begin{Sinput}
grid <- tibble::as_tibble(
  expand.grid(
    hex_long = seq(bbox$min[1] - buffer_dist, # minimum
      bbox$max[1] + buffer_dist, # maximum
      hex_size), # distance between hexagons
    hex_lat = seq(bbox$min[2] - buffer_dist, # minimum
      bbox$max[2] + buffer_dist, # maximum
      hex_size))) # distance between hexagons
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
grid1 <- ggplot() + geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
    dplyr::distinct() %>%
    dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()

# Shift the longitude of every second latitude to the right to make hex structure
grid <- grid %>%
    dplyr::mutate(hex_long = ifelse(hex_lat %in% shift_lat, hex_long,
        hex_long + (hex_size / 2))) %>%
    dplyr::mutate(id=1:NROW(.)) %>%
    dplyr::mutate(assigned=FALSE)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
grid2 <- ggplot() + geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
\end{Sinput}
\end{Schunk}

Each point is now given an ID.

\begin{Schunk}
\begin{Sinput}
full_grid <- grid <- grid %>%
    mutate(hex_long_int = dense_rank(hex_long)-1,
        hex_lat_int = dense_rank(hex_lat)-1)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
gridExtra::grid.arrange(grid1, grid2, nrow = 1)
\end{Sinput}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-2-1} \caption[Grid points to create a tilegram]{Grid points to create a tilegram.}\label{fig:unnamed-chunk-2}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
nlong <- length(unique(grid$hex_long))
nlat <- length(unique(grid$hex_lat))

centroids <- centroids %>%
  mutate(long_int = round((longitude-min(grid$hex_long))/(max(grid$hex_long)-min(grid$hex_long))*nlong, 0),
          lat_int = round((latitude-min(grid$hex_lat))/(max(grid$hex_lat)-min(grid$hex_lat))*nlat, 0))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Split the centroid data set
centroids <- centroids %>% 
  nest(longitude, latitude) %>% 
  mutate(closest = map(data, closest_focal_point, focal_points = capital_cities %>% filter(points == "Hobart"))) %>% 
  unnest(data, closest) %>% 
  arrange(focal_distance)
\end{Sinput}
\end{Schunk}

@@

\hypertarget{allocate-each-centroids-to-a-hexagon-grid-point}{%
\subsection{Allocate each centroids to a hexagon grid
point}\label{allocate-each-centroids-to-a-hexagon-grid-point}}

@@

\begin{Schunk}
\begin{Sinput}
s_centroids <- centroids %>% arrange(focal_distance)

s_centroids <- split(s_centroids, s_centroids[["focal_distance"]])

# Set up allocation data frame
centroid_allocation <- NULL
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
hex_filter = 10
width = 35
# keep value to reset expanded distances
expand_dist <- hex_filter
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# filter for only the available hex grid points
hex_grid <- grid %>% filter(!assigned)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
filter_dist <- hex_filter*hex_size

# filter grid for avaiable points
centroid1 <- centroids %>% head(1)

flong <- centroid1$longitude
flat <- centroid1$latitude

hex_grid <- hex_grid %>% ungroup() %>%
        filter(flat - filter_dist < hex_lat & hex_lat < flat + filter_dist) %>%
        filter(flong - filter_dist < hex_long & hex_long < flong + filter_dist)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
    hex_grid <- hex_grid %>%
        rowwise %>%
        mutate(
        hex_lat_c = hex_lat - flat,
        hex_long_c = hex_long - flong) %>%
        mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1/2))

        f_angle <- centroid1 %>%
            mutate(atan = atan2(latitude-latitude1,longitude-longitude1),
                angle = (atan*180/pi),
                pangle = ifelse(angle<0, angle +360, angle)) %>% pull()

        hex_grid <- hex_grid %>%
            # create circle of radius: hex_filter
            filter(hyp < filter_dist) %>%
            mutate(
                # geosphere takes a long time
                angle = f_angle,
                angle_plus = (angle + width)%%360,
                angle_minus = (angle - width)%%360,
                atan = atan2(hex_lat_c, hex_long_c),
                hex_angle = (atan*180/pi),
                hex_angle = ifelse(hex_angle<0, hex_angle +360, hex_angle))
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
gridExtra::grid.arrange(buff1, buff2, nrow=1)
\end{Sinput}

@@

\begin{Schunk}
\begin{Sinput}
# Filter for angle within circle

        if (hex_grid$angle_minus[1] < hex_grid$angle_plus[1]) {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(angle_minus < hex_angle & hex_angle < angle_plus)
        } else {
            hex_grid <- hex_grid %>%
                # create slice of 60 degrees from centroid
                filter(hex_angle < angle_plus | angle_minus > hex_angle)
        }
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
ggplot() + geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga, fill = "grey", size = 0.3, colour = "lightgrey") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75) + 
  geom_point(aes(x=longitude, y = latitude), data = centroid1, colour = "#b2df8a", size = 0.8) 
\end{Sinput}

\includegraphics{algorithmRjournal_files/figure-latex/plot_slice-1} \end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Choose first available point
cent <- centroid1 %>% dplyr::rename(focal_point = points, focal_dist = focal_distance, focal_angle = angle)

# Filter should give one hex point
hex <- hex_grid %>% 
  ungroup %>% 
  filter(hyp == min(hyp)) %>%
  select(hex_long, hex_lat, hex_id = id)

# Update grid to show this centroid as assigned
hex_grid[which(hex_grid$id == hex$hex_id),]$assigned <- TRUE

# Add to table of allocated centroids
centroid_allocation <- bind_rows(centroid_allocation, dplyr::bind_cols(cent, hex)) %>% as_tibble()
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
hex_points_df <- centroid_allocation %>% 
  fortify_hexagon(hex_size = hex_size,
  sf_id = "LGA_CODE16")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(LGA_NAME16, polygon)), data = fort_lga %>% filter(LGA_CODE16=="62810"), fill = "grey", size = 0.3, colour = "lightgrey") +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75) + 
  geom_point(aes(x=longitude1, y = latitude1), data = centroid1, colour = "#fb9a99", size = 1) + 
  geom_point(aes(x=longitude, y = latitude), data = centroid1, colour = "#b2df8a", size = 0.8) + 
  theme_void() +
  coord_equal()
\end{Sinput}

\includegraphics{algorithmRjournal_files/figure-latex/plot_allocated-1} \end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
\end{Sinput}
\begin{Soutput}
#> Warning in st_centroid.sf(.): st_centroid assumes attributes are constant
#> over geometries of x
\end{Soutput}
\begin{Soutput}
#> Warning in st_centroid.sfc(st_geometry(x), of_largest_polygon =
#> of_largest_polygon): st_centroid does not give correct centroids for
#> longitude/latitude data
\end{Soutput}
\begin{Sinput}
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.3,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.3,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 35,
  verbose = FALSE
)
\end{Sinput}
\begin{Soutput}
#> Allocating centroids, in order of distance to closest focal point.
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
h1 <- hex_allocated %>%
  fortify_hexagon(hex_size = 0.3, 
  sf_id = "LGA_CODE16") %>% 
  left_join(., tas_lga)
\end{Sinput}
\begin{Soutput}
#> Joining, by = "LGA_CODE16"
\end{Soutput}
\begin{Soutput}
#> Warning: Column `LGA_CODE16` joining character vector and factor, coercing
#> into character vector
\end{Soutput}
\begin{Sinput}
p1 <- fortify_sfc(tas_lga)
end_hex <- ggplot() +
  geom_polygon(data = p1, aes(x=long, lat, group = interaction(LGA_NAME16, polygon)), fill = "grey", size = 0.3, colour = "lightgrey") +
  geom_polygon(data = h1, aes(x=long, lat, group = LGA_CODE16)) + theme_void() + coord_equal()

gridExtra::grid.arrange(cents, end_hex, nrow = 1)
\end{Sinput}

>>>>

There are four tasks to create a hexagon tilegram. These steps can be
executed by the main function, \texttt{create\_hexmap}, or can be
implemented separately for more flexibility.

@@

\begin{Schunk}
\begin{Sinput}
# To install sugarbag:
# devtools::install_github("srkobakian/sugarbag")
library(sugarbag)
library(tidyverse)
\end{Sinput}
\end{Schunk}

@@

The \texttt{create\_hexmap} function requires several parameters, if
they are not provided, the information will be derived from the
\CRANpkg{sf} set of shapes used. If users already have a grid and set of
centroids they may choose to only use the \texttt{allocate} function.

\emph{The following must be provided to \texttt{create\_hexmap}:}

@@

\begin{Schunk}
\begin{Sinput}
centroids <- create_centroids(shp_sf = tas_lga, sf_id = "LGA_CODE16")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/end_cents-1} \caption[Local Government Areas and derived centroids]{Local Government Areas and derived centroids.}\label{fig:end_cents}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)
\end{Sinput}
\end{Schunk}

@@

A sequence is created for longitude columns, and another for latitude
rows. The sequences begin at the minimum longitude (latitude), minus the
buffer distance. Equally spaced intervals the size of the hexagons are
created up to the maximum longitude (latitude), plus the buffer
distance. An individual point is created from all intersections of the
longitude columns and latitude row sequences.

@@

\begin{Schunk}
\begin{figure}
\includegraphics{algorithmRjournal_files/figure-latex/unnamed-chunk-3-1} \caption[Grid points to create a tilegram]{Grid points to create a tilegram.}\label{fig:unnamed-chunk-3}
\end{figure}
\end{Schunk}

@@

@@

@@

\hypertarget{allocate-each-centroid-to-a-hexagon-grid-point}{%
\subsection{Allocate each centroid to a hexagon grid
point}\label{allocate-each-centroid-to-a-hexagon-grid-point}}

@@

@@

@@

@@

@@

@@

@@

\includegraphics{algorithmRjournal_files/figure-latex/buffs-1} \end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
# Create centroids set
centroids <- create_centroids(tas_lga, "LGA_CODE16")
# Create hexagon location grid
grid <- create_grid(centroids = centroids,
    hex_size = 0.3,
    buffer_dist = 1.2)
# Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "LGA_CODE16",
  # same column used in create_centroids
  hex_size = 0.3,
  # same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities %>% filter(points == "Hobart"),
  width = 35,
  verbose = FALSE
)
\end{Sinput}
\begin{Soutput}
#> Allocating centroids, in order of distance to closest focal point.
\end{Soutput}
\end{Schunk}

\hypertarget{transform-the-data-for-plotting}{%
\section{Transform the data for
plotting}\label{transform-the-data-for-plotting}}

To plot polygons using \CRANpkg{ggplot} the \CRANpkg{sf} object can be
transformed to a set of indidivual points that can be joined as a
polygon.

\begin{Schunk}
\begin{Sinput}
h1 <- hex_allocated %>%
  fortify_hexagon(hex_size = 0.3, 
  sf_id = "LGA_CODE16") %>% 
  left_join(., tas_lga)
\end{Sinput}
\end{Schunk}
